---
title: "02_statistical_analysis"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
df <- read.csv("../data/preprocessed/processed_appstore.csv")

df$Category <- as.factor(df$Category)
df$Type <- as.factor(df$Type)
```

## Q1: How do install distributions vary across different application categories?

### Method

A boxplot is used to visualize the distribution of install counts
(log-transformed) across different application categories. This helps
compare medians, spread, and variability between categories.

```{r}

par(mar=c(12,4,4,2))
boxplot(log_installs ~ Category, data=df, las=2,
        main="Installs across Categories",
        ylab="Log Installs")
```

### Result

The boxplot indicates moderate variation in install distributions across
categories. Most categories have median log install values in a similar
range (approximately 10–14), with a few categories showing slightly
higher medians and some slightly lower. Additionally, several categories
exhibit wide interquartile ranges and numerous outliers, indicating
variability in app performance within categories.

### Interpretation

The results suggest that while category does influence install
performance, the differences across categories are not extremely large.
Most categories show comparable levels of installs, indicating that
success is not solely dependent on category. However, variability within
each category suggests that individual app characteristics (such as
rating, reviews, and features) also play a significant role. The
presence of outliers further indicates that a few apps achieve
exceptionally high installs regardless of category.

## Q3: Which categories consistently achieve significantly higher install counts?

### Method

ANOVA (Analysis of Variance) is used to determine whether there are
statistically significant differences in install counts across
application categories. Additionally, mean install values for each
category are computed for comparison.

```{r}
aggregate(log_installs ~ Category, data=df, mean)

anova_model <- aov(log_installs ~ Category, data=df)
summary(anova_model)
```

### Result

The ANOVA test shows a highly significant difference in install counts
across application categories (F = 24.73, p \< 2e-16). This indicates
that the mean install counts are not equal across all categories.

From the mean values, categories such as Entertainment (14.37), Game
(13.30), and Education (13.12) have higher average install counts. In
contrast, categories like Medical (8.59), Business (9.38), and Events
(9.48) show lower mean install values.

### Interpretation

The results indicate that application category significantly influences
install performance. Categories related to entertainment and gaming tend
to attract more users due to higher engagement and broader appeal,
resulting in higher install counts. On the other hand, categories such
as Medical and Business cater to more specific user needs, leading to
lower install volumes.

Since the p-value is much smaller than 0.05, the null hypothesis that
all categories have equal mean installs is rejected. This confirms that
category is an important factor affecting app success.

## Q4: How do install patterns differ between Free and Paid applications?

### Method

A t-test is used to compare the mean install counts between Free and
Paid apps.

```{r}
aggregate(log_installs ~ Type, data=df, mean)

t.test(log_installs ~ Type, data=df)
```

### Result

The mean log install count for Free applications is 11.58, while for
Paid applications it is 8.08, showing a large difference between the two
groups.

The Welch two-sample t-test indicates that this difference is highly
statistically significant (t = 27.99, df ≈ 698.52, p \< 2.2e-16). The
95% confidence interval for the difference in means is [3.26, 3.75],
which does not include zero.

### Interpretation

Free applications achieve significantly higher install counts than Paid
applications. The large gap in mean values and the extremely small
p-value indicate that pricing model strongly influences user adoption.

Since p \< 0.05, the null hypothesis (equal mean installs for Free and
Paid apps) is rejected. This suggests that users are much more likely to
download Free apps, making the free pricing model a key driver of
install performance.

## Q5: Does price significantly impact install volume after controlling for rating and category?

### Method

A multiple linear regression model is used to analyze the effect of
price on install counts while controlling for other important variables
such as rating and category.

In this model: - The dependent variable is log-transformed install count
(log_installs). - The independent variables include Price, Rating, and
Category.

This approach allows us to isolate the effect of price on installs while
accounting for differences in app quality (rating) and app type
(category).

```{r}
model_price <- lm(log_installs ~ Price + Rating + Category, data=df)
summary(model_price)
```

### Result

The multiple linear regression model shows that Price has a negative and
statistically significant effect on install counts (coefficient =
-0.00813, p = 0.000289). This indicates that as price increases, install
counts decrease, even after controlling for Rating and Category.

Rating has a positive and significant effect (coefficient = 0.3096, p \<
0.001), suggesting that higher-rated apps tend to receive more installs.

Several categories also show significant effects compared to the
baseline category. For example, Entertainment (3.18, p \< 0.001), Game
(2.09, p \< 0.001), and Education (1.87, p = 0.001) have strong positive
impacts on installs, while categories such as Medical (-2.59, p \<
0.001) and Business (-1.79, p \< 0.001) have negative impacts.

The overall model is statistically significant (F = 24.3, p \< 2.2e-16),
although the R-squared value (0.106) indicates that only about 10.6% of
the variation in installs is explained by the model.

### Interpretation

The results indicate that price significantly affects install
performance, even after accounting for rating and category. The negative
coefficient for price confirms that higher-priced apps tend to attract
fewer users, suggesting that cost acts as a barrier to downloads.

At the same time, rating plays a strong positive role, indicating that
users are more likely to install apps with better user feedback.
Category also influences installs, with entertainment and gaming-related
apps performing better due to higher engagement, while specialized
categories like Medical perform worse.

Although the model is statistically significant, the relatively low
R-squared value suggests that other factors (such as marketing, app
quality, or user experience) also contribute to install success.

Overall, the findings show that price has a statistically significant
but relatively small negative effect on installs when other variables
are controlled.

## Q6: What is the statistical strength of correlation between user ratings and install levels?

### Method

Pearson’s correlation test is used to measure the strength and direction
of the linear relationship between user ratings and install counts
(log-transformed). This test provides both the correlation coefficient
and the statistical significance of the relationship.

```{r}
cor.test(df$Rating, df$log_installs)
```

### Result

The Pearson correlation coefficient between rating and log install count
is 0.0597, indicating a very weak positive relationship.

The correlation is statistically significant (t = 5.02, p = 5.43e-07),
and the 95% confidence interval [0.036, 0.083] does not include zero.

### Interpretation

The results indicate that while there is a statistically significant
relationship between rating and install count, the strength of the
relationship is very weak.

This means that higher-rated apps tend to have slightly more installs,
but rating alone does not strongly influence install performance. Other
factors such as category, price, and user engagement are likely more
important in determining app success.

## Q7: Does review count independently influence install performance?

### Method

A multiple linear regression model is used to evaluate the impact of
review count on install performance while controlling for other
important variables such as rating and category.

In this model: - The dependent variable is log-transformed install count
(log_installs). - The independent variables include Reviews, Rating, and
Category.

This approach allows us to determine whether review count has an
independent effect on installs after accounting for app quality (rating)
and differences across categories.

```{r}
model_reviews <- lm(log_installs ~ Reviews + Rating + Category, data=df)
summary(model_reviews)
```

### Result

The regression model shows that Reviews has a positive and highly
significant effect on install counts (coefficient = 7.43e-07, p \<
2e-16). This indicates that apps with more reviews tend to have higher
install counts.

Rating also remains a significant predictor (coefficient = 0.229, p =
0.0015), showing that higher-rated apps are associated with more
installs.

Several categories continue to show significant effects, such as
Entertainment (3.06, p \< 0.001), Education (1.85, p \< 0.001), and Game
(1.62, p \< 0.001), while categories like Medical (-2.61, p \< 0.001)
show lower installs.

The overall model is statistically significant (F = 35.56, p \<
2.2e-16), and the R-squared value is 0.147, indicating that
approximately 14.7% of the variation in install counts is explained by
the model.

### Interpretation

The results indicate that review count has a strong and independent
influence on install performance. Apps with more reviews tend to attract
more users, likely because higher review counts signal popularity and
trustworthiness.

Even after controlling for rating and category, reviews remain highly
significant, suggesting that user engagement plays a crucial role in app
success. While the coefficient value is small due to the scale of
reviews, its strong statistical significance highlights its importance.

The higher R-squared compared to previous models suggests that including
reviews improves the model’s explanatory power. Overall, review count is
a key predictor of install performance, reflecting the importance of
user feedback and social proof in driving app downloads.

## Q8: Is there a rating threshold above which install probability increases significantly?

### Method

Two approaches are used to analyze the effect of rating on install
success:

1.  Logistic regression is applied with high_success (whether installs
    exceed a threshold, e.g., 1M+) as the dependent variable and Rating
    as the predictor. This helps estimate how rating affects the
    probability of achieving high installs.

2.  Rating values are grouped into bins (e.g., 0–3, 3–4, 4–4.5, 4.5–5),
    and the average success rate within each bin is calculated. This
    helps identify any threshold effect visually.

```{r}
logit_model <- glm(high_success ~ Rating, data=df, family="binomial")
summary(logit_model)
```

```{r}
df$rating_bin <- cut(df$Rating, breaks=c(0,3,4,4.5,5))

aggregate(high_success ~ rating_bin, data=df, mean)
```

### Result

The logistic regression model shows that Rating has a positive and
statistically significant effect on the probability of achieving high
install counts.

The binned analysis indicates that the proportion of high-success apps
increases as rating increases, with higher rating bins (especially above
4.0 or 4.5) showing a greater likelihood of achieving high installs.

### Interpretation

The results suggest that higher-rated apps are more likely to achieve
high install milestones. This indicates the presence of a rating
threshold, where apps above a certain rating level (around 4.0–4.5)
experience a noticeable increase in success probability.

However, the increase is gradual rather than abrupt, indicating that
rating contributes to success but is not the sole determining factor.
Other variables such as reviews, category, and pricing also play
important roles in influencing install performance.

## Q9: Are Free applications statistically more likely to achieve high install milestones (e.g., 1M+) than Paid applications?

### Method

A chi-square test of independence is used to examine the relationship
between app type (Free vs Paid) and install success (high_success). This
test determines whether the likelihood of achieving high installs
depends on the pricing model.

```{r}
table(df$Type, df$high_success)

chisq.test(table(df$Type, df$high_success))
```

### Result

The contingency table shows that a large number of Free apps achieve
high installs (2468) compared to Paid apps (only 19).

The chi-square test indicates a highly significant association between
app type and install success (χ² = 259.08, df = 1, p \< 2.2e-16).

### Interpretation

The results clearly show that Free applications are significantly more
likely to achieve high install milestones compared to Paid applications.
The extremely small p-value leads to rejection of the null hypothesis,
confirming that install success is dependent on app type.

This suggests that the free pricing model strongly increases the
probability of reaching high install counts, likely because it removes
cost barriers and encourages wider adoption among users.

## Q10: Is there an optimal price range for paid applications that maximizes install performance?

### Method

Prices are grouped into ranges using binning, and the average install
counts (log-transformed) are calculated for each price range. This helps
identify patterns in install performance across different pricing
levels.

```{r}
df$price_bin <- cut(df$Price, breaks=c(0,1,5,10,50))

aggregate(log_installs ~ price_bin, data=df, mean)
```

### Result

The average log install counts for different price ranges are as
follows:

(0,1] → 7.54 (1,5] → 8.25 (5,10] → 8.36 (10,50] → 7.52

The mid-range prices (1–10) show slightly higher install values compared
to very low (0–1) and very high (10–50) price ranges.

### Interpretation

The results suggest that extremely low-priced and high-priced apps tend
to have lower install counts, while moderate price ranges (1–10) perform
slightly better.

This indicates that there may be a mild optimal pricing range where
users perceive better value, leading to higher installs. However, the
differences across price ranges are not very large, suggesting that
price alone is not a dominant factor in determining install performance.

Overall, pricing has a moderate influence, and other factors such as
category, ratings, and reviews likely play a more significant role in
app success.

## Q11: Does app size significantly influence installs or ratings?

### Method

Pearson’s correlation test and a simple linear regression model are used
to examine the relationship between app size and install counts
(log-transformed).

-   Correlation analysis measures the strength and direction of the
    relationship.
-   Linear regression evaluates whether app size significantly predicts
    install performance.

```{r}
cor.test(df$Size, df$log_installs)

model_size <- lm(log_installs ~ Size, data=df)
summary(model_size)
```

### Result

The correlation between app size and log install count is -0.00998,
indicating an extremely weak relationship. The result is not
statistically significant (p = 0.4027), and the 95% confidence interval
[-0.033, 0.013] includes zero.

Similarly, the regression model shows that Size is not a significant
predictor of install counts (coefficient = -0.000367, p = 0.403). The
R-squared value is approximately 0.0001, indicating that app size
explains almost none of the variation in installs.

### Interpretation

The results indicate that app size has no meaningful impact on install
performance. The extremely low correlation and non-significant p-value
suggest that users do not consider app size as an important factor when
deciding to install an app.

This implies that other factors such as category, price, ratings, and
reviews play a much more important role in determining app success.
Therefore, app size is not a useful predictor of install performance.

## Q12: Do recently updated applications outperform older applications in install counts?

### Method

A linear regression model is used to analyze the relationship between
app update recency and install counts.

-   The variable days_since_update is calculated as the number of days
    between the current date and the last update date.
-   The dependent variable is log-transformed install count
    (log_installs).
-   This model helps determine whether more recently updated apps
    achieve higher installs.

```{r}
df$Last_Updated <- as.Date(df$Last_Updated)
df$days_since_update <- as.numeric(Sys.Date() - df$Last_Updated)

model_update <- lm(log_installs ~ days_since_update, data=df)
summary(model_update)
```

### Result

The regression model shows that days_since_update has a negative and
statistically significant effect on install counts (coefficient =
-0.00164, p \< 2e-16).

This indicates that as the number of days since the last update
increases, install counts decrease.

The model is statistically significant (F = 276.4, p \< 2.2e-16),
although the R-squared value (0.038) suggests that only about 3.8% of
the variation in installs is explained by update recency.

### Interpretation

The results indicate that recently updated applications tend to achieve
higher install counts compared to older applications. The negative
relationship suggests that apps become less attractive to users as they
remain outdated.

Frequent updates may signal active development, improved features, and
better user experience, which encourages more downloads.

However, the relatively low R-squared value indicates that while update
recency is significant, it is not the primary factor influencing install
performance. Other variables such as category, pricing, ratings, and
reviews play a more substantial role.

Overall, keeping apps updated is an important factor for maintaining
user interest and improving install performance.

```{r}

```
