---
output:
  pdf_document: default
  html_document: default
---
```r
---
title: "App Store Success"
date: "2026-04-10"
output: pdf_document
---

```{r setup, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
	library(dplyr)
	library(readr)
	library(stringr)
	library(ggplot2)
})
set.seed(42)
options(width = 120)
```

### Paths

```{r}
raw_apps_path <- "data/raw/googleplaystore.csv"
raw_reviews_path <- "data/raw/googleplaystore_user_reviews.csv"

clean_apps_path <- "data/cleaned/apps_cleaned.csv"
review_summary_path <- "data/cleaned/review_summary.csv"
final_data_path <- "data/cleaned/final_data.csv"
final_filtered_path <- "data/cleaned/final_data_filtered.csv"
processed_path <- "processed_appstore.csv"

output_dir <- "outputs"
plot_dir <- file.path(output_dir, "plots")
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
if (!dir.exists(plot_dir)) dir.create(plot_dir, recursive = TRUE)
```

### Load Raw Data

```{r}
apps <- read_csv(raw_apps_path, show_col_types = FALSE)
reviews <- read_csv(raw_reviews_path, show_col_types = FALSE)
head(apps)
head(reviews)
```

### Clean Apps Data

```{r}
apps_clean <- apps %>%
	filter(!is.na(Category), !is.na(Rating)) %>%
	distinct(App, .keep_all = TRUE) %>%
	mutate(
		Reviews = as.numeric(Reviews),
		Installs = as.numeric(str_replace_all(str_replace_all(Installs, "\\+", ""), ",", "")),
		Price = as.numeric(str_replace(Price, "\\$", "")),
		Size = ifelse(Size == "Varies with device", NA, Size),
		Size = str_replace(Size, "k", ""),
		Size = str_replace(Size, "M", ""),
		Size = as.numeric(Size)
	)

apps_clean <- apps_clean %>%
	mutate(
		Category = as.factor(Category),
		Type = as.factor(Type),
		`Content Rating` = as.factor(`Content Rating`),
		Genres = as.factor(Genres)
	)

write_csv(apps_clean, clean_apps_path)
knitr::kable(head(apps_clean), caption = "Cleaned Apps Data")
```

### Clean Reviews and Summarize

```{r}
reviews_clean <- reviews %>%
	filter(!is.na(Translated_Review)) %>%
	select(App, Sentiment_Polarity)

reviews_clean$Sentiment_Polarity[is.na(reviews_clean$Sentiment_Polarity)] <- 0

review_summary <- reviews_clean %>%
	group_by(App) %>%
	summarise(
		Avg_Sentiment = mean(Sentiment_Polarity, na.rm = TRUE),
		Review_Count = n(),
		.groups = "drop"
	)

write_csv(review_summary, review_summary_path)
knitr::kable(head(review_summary), caption = "Review Summary")
```

### Merge Data

```{r}
final_data <- apps_clean %>%
	left_join(review_summary, by = "App") %>%
	mutate(
		Avg_Sentiment = ifelse(is.na(Avg_Sentiment), 0, Avg_Sentiment),
		Review_Count = ifelse(is.na(Review_Count), 0, Review_Count),
		Success = ifelse(Installs >= 1000000, 1, 0),
		Success = as.factor(Success)
	)

write_csv(final_data, final_data_path)

final_data_filtered <- final_data %>% filter(!is.na(Success))
write_csv(final_data_filtered, final_filtered_path)

knitr::kable(head(final_data), caption = "Merged Final Data")
```

### Build Processed Dataset

```{r}
df <- final_data

df <- df %>%
	mutate(
		Installs = as.numeric(Installs),
		Reviews = as.numeric(Reviews),
		Rating = as.numeric(Rating),
		Price = as.numeric(Price),
		Size = as.numeric(Size),
		log_installs = log1p(Installs),
		log_reviews = log1p(Reviews),
		is_free = ifelse(Type == "Free", 1, 0),
		high_success = ifelse(Installs >= 1000000, 1, 0),
		Last_Updated = as.Date(`Last Updated`, format = "%B %d, %Y"),
		days_since_update = as.numeric(Sys.Date() - Last_Updated)
	)

if ("Genres" %in% names(df)) df$Genres <- NULL
names(df)[names(df) == "Content Rating"] <- "Content.Rating"

df <- df %>%
	mutate(
		Category = as.factor(Category),
		Type = as.factor(Type),
		Content.Rating = as.factor(Content.Rating)
	) %>%
	na.omit()

write.csv(df, processed_path, row.names = FALSE)

cat("Rows in final dataset:", nrow(df), "\n")
```

### Plot Helper

```{r}
save_plot <- function(plot_obj, filename, width = 10, height = 6) {
	ggsave(filename = file.path(plot_dir, filename), plot = plot_obj, width = width, height = height)
}

base_plot <- theme_minimal(base_size = 12) +
	theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())
```

### Q1

```{r}
q1_summary <- df %>% group_by(Category) %>% summarise(mean_log_installs = mean(log_installs), .groups = "drop")
knitr::kable(q1_summary)

p1 <- ggplot(df, aes(x = reorder(Category, log_installs, FUN = median), y = log_installs)) +
	geom_boxplot(fill = "#4C78A8") + coord_flip() + labs(title = "Q1") + base_plot
print(p1)
```

### Q2

```{r}
q2_top_categories <- q1_summary %>% arrange(desc(mean_log_installs)) %>% slice_head(n = 10)
knitr::kable(q2_top_categories)

p2 <- ggplot(q2_top_categories, aes(x = reorder(Category, mean_log_installs), y = mean_log_installs)) +
	geom_col(fill = "#F58518") + coord_flip() + labs(title = "Q2") + base_plot
print(p2)
```

### Q3

```{r}
q3_anova <- aov(log_installs ~ Category, data = df)
summary(q3_anova)

p3 <- ggplot(df, aes(x = Category, y = log_installs)) +
	geom_violin() + coord_flip() + labs(title = "Q3") + base_plot
print(p3)
```

### Q4

```{r}
q4_type_means <- aggregate(log_installs ~ Type, data = df, mean)
knitr::kable(q4_type_means)

q4_ttest <- t.test(log_installs ~ Type, data = df)
q4_ttest

p4 <- ggplot(df, aes(x = Type, y = log_installs)) +
	geom_boxplot() + labs(title = "Q4") + base_plot
print(p4)
```

### Q5

```{r}
q5_model_price <- lm(log_installs ~ Price + Rating + Category, data = df)
summary(q5_model_price)

p5 <- ggplot(df, aes(x = Price, y = log_installs)) +
	geom_point() + geom_smooth(method = "lm") + labs(title = "Q5") + base_plot
print(p5)
```

### Q6

```{r}
q6_cor <- cor.test(df$Rating, df$log_installs)
q6_cor

p6 <- ggplot(df, aes(x = Rating, y = log_installs)) +
	geom_point() + geom_smooth(method = "lm") + labs(title = "Q6") + base_plot
print(p6)
```

### Q7

```{r}
q7_model_reviews <- lm(log_installs ~ Reviews + Rating + Category, data = df)
summary(q7_model_reviews)

p7 <- ggplot(df, aes(x = log_reviews, y = log_installs)) +
	geom_point() + geom_smooth(method = "lm") + labs(title = "Q7") + base_plot
print(p7)
```

### Q8

```{r}
q8_logit <- glm(high_success ~ Rating, data = df, family = "binomial")
summary(q8_logit)

df$rating_bin <- cut(df$Rating, breaks = c(0,3,4,4.5,5), include.lowest = TRUE)
q8_bin_success <- aggregate(high_success ~ rating_bin, data = df, mean)
knitr::kable(q8_bin_success)

p8 <- ggplot(q8_bin_success, aes(x = rating_bin, y = high_success)) +
	geom_line() + geom_point() + labs(title = "Q8") + base_plot
print(p8)
```

### Q9

```{r}
q9_table <- table(df$Type, df$high_success)
q9_chi <- chisq.test(q9_table)
q9_chi

q9_df <- as.data.frame(q9_table)
knitr::kable(q9_df)

p9 <- ggplot(q9_df, aes(x = Var1, y = Freq, fill = Var2)) +
	geom_col(position="dodge") + labs(title = "Q9") + base_plot
print(p9)
```

### Q10

```{r}
paid_df <- df %>% filter(Type == "Paid", Price > 0)
paid_df$price_bin <- cut(paid_df$Price, breaks = c(0,1,5,10,50), include.lowest = TRUE)
q10_price_range <- aggregate(log_installs ~ price_bin, data = paid_df, mean)
knitr::kable(q10_price_range)

p10 <- ggplot(q10_price_range, aes(x = price_bin, y = log_installs)) +
	geom_col() + labs(title = "Q10") + base_plot
print(p10)
```

### Q11

```{r}
cor.test(df$Size, df$log_installs)
summary(lm(log_installs ~ Size, data = df))
```

### Q12

```{r}
summary(lm(log_installs ~ days_since_update, data = df))
```

### Q13

```{r}
q13_category_success <- aggregate(high_success ~ Category, data = df, mean)
knitr::kable(q13_category_success)
```

### Q14

```{r}
q14_model <- lm(log_installs ~ Category + log_reviews + Rating + Price + Size + Avg_Sentiment + days_since_update, data = df)
q14_summary <- summary(q14_model)
q14_summary
```

### Q15

```{r}
df$Success <- as.factor(ifelse(df$Installs >= 1000000, 1, 0))

train_idx <- sample(seq_len(nrow(df)), size = floor(0.8 * nrow(df)))
train <- df[train_idx, ]
test <- df[-train_idx, ]

model <- glm(Success ~ Category + Rating + Reviews + Price + Size + Avg_Sentiment + days_since_update, data = train, family="binomial")

pred <- predict(model, newdata=test, type="response")
pred_class <- ifelse(pred>0.5,1,0)

conf <- table(pred_class, test$Success)

acc <- mean(pred_class == as.numeric(as.character(test$Success)))
cat("Accuracy:", acc)
```

```
```
